---
title: "Parameter Estmination: Wild Fish Catch"
author: "Teague Tran"
date: "3/5/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, message = FALSE, warning = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, message = FALSE, warning = FALSE}
#set up
library(tidyverse)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)
library(janitor)

#add in csv and change column names to name of first row 
#units = million tons
fish_catch <- read_csv(here("data", "fish_catch.csv")) %>% 
  drop_na() %>% 
  row_to_names(row_number = 1) %>% 
  clean_names()
#change year to have 1950 start at 0
fish_catch <- fish_catch %>% 
  mutate(year_num = as.numeric(year)) %>% 
  mutate(wild_catch = as.numeric(wild_catch, replace = TRUE)) %>% 
  mutate(year_new = year_num - min(year_num)) %>% 
  select(wild_catch, year_new)
```

```{r, message = FALSE, warning = FALSE}
# visualize the data 
ggplot( data = fish_catch, aes( x = year_new, y = wild_catch)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "Year", y = "Wild fish catch (million tons)") 
#look at log transformed data: 

ggplot( data = fish_catch, aes( x = year_new, y = log(as.numeric(wild_catch)))) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "Year", y = "ln(wild fish catch) (million tons)") 
#K looks to be at around 90 million tons
```


The log-transformed graph shows evidence of a logistic growth curve. Therefore, we will be using the equation below for the model. 

$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant; equivalent to $r$

After further analysis, the following are the initial values for the parameters in the model :

- $K$ = 90
- $A$ = 4.23
- $k$ = .0354

```{r, message = FALSE, warning = FALSE}
#try to find growth rate constant (k) during exponential growth phase
fish_catch_exp <- fish_catch  %>% 
  filter(year_new <= 40) %>% 
  mutate(ln_prod = log(as.numeric(wild_catch)))

#find k estimate using slope of linear equation
lm_k <- lm(ln_prod ~ year_new, data = fish_catch_exp)

#find A; K at approximately 95 million
#A = (90-17.2)/17.2 = 4.23

#perform nls 
df_nls <- nls(as.numeric(wild_catch) ~ K/(1 + A*exp(-r*year_new)),
              data = fish_catch,
              start = list(K = 90, A = 4.23, r = .03255),
              trace = FALSE)

```


Our model with estimated parameters is:
$$P(t) = \frac{90}{1+4.23e^{-.0354t}}$$
```{r, message = FALSE, warning = FALSE}
# visualize model over original observed values
p_predict <- predict(df_nls)

# Bind predictions to original data frame: 
fish_catch_complete <- data.frame(fish_catch, p_predict)

#plot both prediction and original data together
ggplot(data = fish_catch_complete, aes(x = year_new, y = wild_catch, colour = "Observed")) +
  geom_point(aes(shape = "Observed fish catch"), color = "orange") + 
  geom_line(aes(x = year_new, y = p_predict, linetype = "Model Prediction"), color = "black",  size = 1) +
  theme_minimal() +
  labs(y = "Wild fish count (million tons)", x = "Year(s) from 1950", title = "Observed Fish Catch vs Fish Catch Calculated by a \nLogicstic Growth Model", shape = "", linetype = "")

# find confidence intervals
df_ci <- confint2(df_nls)
```

## Citation:

 Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.
